Dirac neutrino mass generation from Majorana messenger

### 2019-09-14 draft.tex
Updates from Overleaf

\section{Dark matter}
\label{sec:DM}
From the charge assignment presented in Table~\ref{tab:partcont} we have that a residuar $Z_2$ symmetry is left over after the $\operatorname{U}(1)_X$ symmetry breaking, with the particles circulating the one-loop neutrino mass diagram  (see Fig.~\ref{fig:zee}) and $\nu_{R3}$ being odd under it whereas $\nu_{R1}, \nu_{R2}$ and and all the SM particles are even. Thus the lightest neutral $Z_2$-odd particle can be a dark matter candidate. In other words, this model features either fermion ($\psi_1$) or scalar ($\chi_{1R}$ or $\chi_{1I}$) dark matter\footnote{Note that $Z'$ is not stable because the $Z'\to\bar{\nu}_{R3}\nu_{R3}$ decay channel can not be kinematically closed.}.

%\subsection{Fermion dark matter}
Since $\psi_{R1}$ is a singlet under the SM gauge group its thermal relic density is controlled by the Yukawa and $\operatorname{U}(1)_X$ gauge interactions. 
The scenario where $\psi_{1}$ self-annihilates dominantly through $h_{i\alpha}$-mediated interactions resembles the very well known scotogenic model \cite{Ma:2006km}. 
In this case sizable $h_{i\alpha}$ are required to reproduce the correct DM abundance, which in turn leads to a mild tension with experimental upper bounds on the rates for rare charged lepton decays. 
The phenomenology of this scenario has been extensively studied  \cite{Kubo:2006yx,Sierra:2008wj,Suematsu:2009ww,Gelmini:2009xd,Hambye:2009pw,Aoki:2010tf,Schmidt:2012yg,Ma:2012if,Kashiwase:2012xd,Kashiwase:2013uy,Racker:2013lua,Toma:2013zsa,Molinaro:2014lfa,Vicente:2014wga,Merle:2015gea,Merle:2015ica,Merle:2016scw,Hagedorn:2016dze,Lindner:2016kqk,Hessler:2016kwm}.
On the other hand, when the $Z'$ portal is the main gate to visible sector, $\psi_1$ largely annihilates  into neutrinos (see~\cite{Blanco:2019hah} for a recent review on $Z'$-mediated Majorana DM). 
% check Ibarra papers
%also through the annihilation
%$\psi_{1} \psi_{1} \to l _{\alpha}^+ \overline{l}_{\beta}^- $, via
%Yukawa coupling $h^{a}_{i}$. These Yukawa couplings are restricted by
%CLFV and neutrino masses. %But, Yukawa couplings $y^{a}_{j}$ is free.
%In the case where the mass of the other fermions and scalars is nearly
%degenerated to $\psi_{1} $, the processes of coannihilation can be
%relevant for the relic density calculation~\cite{Klasen:2013jpa}.

@@

%\subsection{Scalar dark matter}
Contrary to fermionic DM, scalar DM has additional interactions beyond the Yukawa couplings  which can be used to control relic density. This corresponds to the simplified dark matter model of scalar singlet-doblet dark matter, which have been studied in detail in~\cite{Restrepo:2019soi}

When the lightest scalar  is mainly a scalar doublet, the phenomenology will be similar inert doublet model~\cite{Honorez:2010re}. For this kind of DM model  inert doublet model there are two mass regions for DM which can to reproduce adequately the relic density ($50$GeV \textless $m_{DM}$ \textless $70$GeV and $535$GeV \textless $m_{DM}$ \textless $20$TeV)~\cite{Garcia-Cely:2015khw}.

@@

Now, if the lightest scalar is mainly a singlet, the DM will be
similar to Singlet Scalar Dark Matter
Model~\cite{Athron:2017kgt}\todo{change for the original ref by Ma et
  al. etc}.
The mass range for the scalar singlet that produces the adequate relic
density is around the Higgs resonance and for masses above $1$TeV. For
this model~\cite{}, the relevant c

oupling for DM phenomenology is the
Higgs portal trough the Lagrangian

 term $\sigma^{2} H^{2}_{i}$.
This term generates interesting phenomenological consequences.
Such as, thermal production of DM in the early
universe~\cite{Yaguna:2008hd}, direct detection and invisible decay of
SM Higgs through $h \to \sigma \sigma$~\cite{Mambrini:2011ik}.
In addition, this model has important contributions for cosmology, in
phenomena such as inflation~\cite{Lerner:2009xg} and
baryogenesis~\cite{Cline:2012hg}.

\section{Cosmological constraints}
\label{sec:CosmoConstraints}
A not-so-minimal model can be obtained by introducing additional chiral SM singlet fields. 
In this case we can use the solutions to the anomaly cancellation conditions found for an Abelian gauge symmetry $\operatorname{U(1)_X}$ in~\cite{Appelquist:2002mw,Campos:2017dgc,Okada:2018tgy,Jana:2019mez}.
We use $f$ ($f$) to denote the $X$-charge of the field $f_R$ ($F_L$) for one generation of the SM. The three linear anomalies in $\operatorname{U(1)_X}$ allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $\alpha$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \sum_{\alpha} n_{\alpha}+3 (e-2l)=&0\,, &   \sum_{\alpha} n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
There are many known solutions for the case in which $e-2l=1$. The full set of standard model charges in terms of $l$ is just
\begin{align}
  u=&-1-\frac{4l}{3}\,,&d=&1+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&1+2l\,,&h=&-1-l\,.
  \label{Eq:SMCharges}
\end{align}
The same solution but in terms of $h$ is given in ~\cite{Okada:2018tgy}.

@@

@@

\section{General solution}

The general solution to fig.~\ref{fig:zee} is
\begin{align}
  L=&\nu_R+4\psi_R -H \,,& \eta=& H-\nu_R-3\psi_R \nonumber\\
  \sigma=& -\nu_R-\psi_R\,,&S=& 2\psi_R \,.
\end{align}
The consistency condition for $H=0$ is
\begin{align}
  L=\nu_R+4\psi_R\,.
\end{align}

On the other hand, for fig.~\ref{fig:ld5}
\begin{align}
  L=&-H+\nu_R+2\psi_R &
  \eta=&H-\nu_R-\psi_R \nonumber\\
  \sigma=& -\nu_R-\psi_R\,,&S=& 2\psi_R \,.
\end{align}
$\nu_R$, $H$  and $L$ are fixed from the anomaly cancellation conditions.

\section{5d model}
\label{sec:5dmodel}
A not-so-minimal model can be obtained as the one-loop realization of
the effective 5-dimensional operator for Dirac neutrino masses~\cite{Gu:2006dc}
\begin{align}
  \label{eq:ld5}
  \mathcal{L}_{5D}=\frac{1}{\Lambda} \overline{L_i} H \nu_R S\,,
\end{align}
as displayed in fig.~\ref{fig:ld5}.
In this case we can use the solutions to the anomaly cancellation conditions found for an Abelian gauge symmetry $\operatorname{U(1)_X}$ in~\cite{Campos:2017dgc,Okada:2018tgy}.
We use $f$ ($f$) to denote the $X$-charge of the field $f_R$ ($F_L$) for one generation of the SM. The three linear anomalies in $\operatorname{U(1)_X}$ allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $\alpha$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \sum_{\alpha} n_{\alpha}+3 (e-2l)=&0\,, &   \sum_{\alpha} n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
There are many known solutions for the case in which $e-2l=1$. The full set of standard model charges in terms of $l$ is just
\begin{align}
  u=&-1-\frac{4l}{3}\,,&d=&1+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&1+2l\,,&h=&-1-l\,.
  \label{Eq:SMCharges}
\end{align}
The same solution but in terms of $h$ is given in ~\cite{Okada:2018tgy}.

For this kind of solutions we have that
\begin{align}
  \psi=&-\frac{\nu+1}{2}\,,&\eta=&-\frac{\nu+1}{2}-l\,,&\sigma=&\frac{1-\nu}{2}\,.
\end{align}

We are interested here in the case of the exotic set of $X$-charges
charges: $\mp 4$,~$\mp 4$,~$\pm 5$, that when assigned to the
three light right-handed neutrinos can forbid three level Dirac and
Majorana masses~\cite{Calle:2018ovc}. This set is compatible with the
Abelian gauge groups $X=B-L,B,D,G$ defined in~\cite{Campos:2017dgc}.
The presence of a set of heavy Majorana mediators to realize the
dimension-5 operator at one-loop with the diagram in
fig.~\ref{fig:ld5}., would spoil the anomaly cancellation conditions.
However, we can further add an extra set of quiral fermion in a
vector-like way, such that the full set heavy fermions do not affect
the anomaly cancellation. A such assignment is displayed in
Table~\ref{tab:partcont2}.
The fields $\xi_{L \alpha}$ guarantee the anomaly cancellation and
their Majorana masses are obtained in the same way than the ones for
$\psi_{R\alpha}$.

\begin{figure}
\centering
\includegraphics[scale=0.75]{Neutrino2.pdf}
\caption{One-loop realization of the dimension-5 effective Dirac neutrino mass operator }
\label{fig:ld5}
\end{figure}
%

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_B$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$\\ \hline
$L $  & $\boldsymbol{2}$& $-1/2$  &  $l$&  $-1$&    $0$ &  $-3/2$&  $-1/2$ \\    
$d_R $  & $\boldsymbol{1}$& $-1/2$  &  $1+2l/3$&  $1/3$&    $1$&  $0$&  $2/3$ \\
$u_R $  & $\boldsymbol{1}$& $+2/3$  &  $-1-4l/3$&  $1/3$&   $-1$&  $1$&  $-1/3$ \\
$Q $  & $\boldsymbol{2}$& $-1/6$  &  $-l/3$& $1/3$&    $0$&  $1/2$&  $1/6$ \\\hline
$e_R $  & $\boldsymbol{1}$& $-1$  &  $1+2l$&  $-1$&    $1$ &  $-2$&  $0$ \\
$H $  & $\boldsymbol{2}$& $1/2$  &  $-1-l$&  $0$&    $-1$ &  $1/2$&  $-1/2$ \\
$\eta$  & $\boldsymbol{2}$ & $+1/2$  & $3/2-l$& $5/2$& $3/2$ &$3$&$2$ \\
$S$ & $\boldsymbol{1}$ & $0$  & $3$& $3$&  $3$ & $3$& $3$ \\
$\sigma$ & $\boldsymbol{1}$ & $0$ & $5/2$& $5/2$&  $5/2$ & $5/2$& $5/2$ \\
\hline
 $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$& $-4$&  $-4$ & $-4$& $-4$\\
$\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $+5$& $+5$&  $5$ & $5$& $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/2$ & $3/2$&  $3/2$ & $3/2$& $3/2$ \\\hline
$\xi_{L\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/2$& $3/2$ &  $3/2$ & $3/2$& $3/2$\\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. The SM fields have the usual $U(1)_{B-L}$ assignment. Now $\alpha=1,2$}
    \label{tab:partcont2}
\end{table}

In this way we end up with a model with four Majorana fermions, two more than in the minimal solution. Alternatively we can have a single set of $\psi_R$ and $\xi_L$ but two set of scalars $\eta_\alpha$, $\sigma_{\alpha}$~\cite{Reig:2018mdk}. If we choose $r$ as a fraction with denominator $2$, let say $r=3/2$, the $\operatorname{U}(1)_{B-L}$ left out a remnant $Z_2$ discrete symmetry which guarantees the stability of the lightest odd-particle.

The two additional left-handed fields $\xi_{L\alpha}$,  with a $\operatorname{U}(1)_{B-L}$ charge of $r=3/2$,  cancel out the anomalies. Moreover, we have both Dirac mass terms $\overline{\xi_{L \alpha}}\psi_{R \beta}$ and $ \overline{\xi_{L\alpha}^c }\xi_{L \beta} \langle S\rangle$ and the full set of Majorana fields are massive and heavy. The lighter of them a would be dark matter candidate.

>>>>

\section{Dark matter}
\label{sec:DM}
From the model charge assignment in Table~\ref{tab:partcont} we have that a residual $Z_2$ symmetry is left over after the $\operatorname{U}(1)_X$ symmetry breaking, with the particles circulating the one-loop neutrino mass diagram  (see Fig.~\ref{fig:zee}) and $\nu_{R3}$ being odd under it whereas $\nu_{R1}, \nu_{R2}$ and all the SM particles trivially transforming. Thus the lightest electrically-neutral $Z_2$-odd particle becomes a dark matter candidate. In other words, this model also provides a solution to the DM puzzle via either fermion ($\psi_1$) or scalar ($\chi_{1R}$ or $\chi_{1I}$) dark matter\footnote{Note that $Z'$ can not constitute a DM candidate due to the instability associated to the $Z'\to\bar{\nu}_{R3}\nu_{R3}$ decay channel, since it can not be kinematically closed.}.

%\subsection{Fermion dark matter}
Since $\psi_{1}$ is a singlet under the SM gauge group its thermal relic density is controlled by the Yukawa and $\operatorname{U}(1)_X$ gauge interactions. 
The scenario where $\psi_{1}$ self-annihilates dominantly through $h_{i\alpha}$-mediated interactions resembles the very well known scotogenic model \cite{Ma:2006km},  
where sizable $h_{i\alpha}$ are required to reproduce the correct DM abundance, which in turn leads to a mild tension with experimental upper bounds on the rates for rare charged lepton decays~\cite{Kubo:2006yx,Sierra:2008wj,Ibarra:2016dlb}. 
%The phenomenology of this scenario has been extensively studied  \cite{Kubo:2006yx,Sierra:2008wj,Suematsu:2009ww,Gelmini:2009xd,Hambye:2009pw,Aoki:2010tf,Schmidt:2012yg,Ma:2012if,Kashiwase:2012xd,Kashiwase:2013uy,Racker:2013lua,Toma:2013zsa,Molinaro:2014lfa,Vicente:2014wga,Merle:2015gea,Merle:2015ica,Merle:2016scw,Hagedorn:2016dze,Lindner:2016kqk,Hessler:2016kwm,Ibarra:2016dlb}.
On the other hand, when the $Z'$ portal \cite{Langacker:2008yv,Dudas:2009uq} is the main gate to visible sector, $\psi_1$ largely annihilates into neutrinos in such a way the observed DM abundance can be reproduced without entering in conflict with the DM searches, which follows from the fact that the $Z'$ does not couples to quarks neither to charged leptons)
(see~\cite{Dudas:2009uq,Arcadi:2013qia,Alves:2013tqa,Alves:2015pea,Alves:2016cqf,Blanco:2019hah} for related phenomenological studies on $Z'$-mediated Majorana DM). 
% check Ibarra papers
%also through the annihilation
%$\psi_{1} \psi_{1} \to l _{\alpha}^+ \overline{l}_{\beta}^- $, via 
%Yukawa coupling $h^{a}_{i}$. These Yukawa couplings are restricted by
%CLFV and neutrino masses. %But, Yukawa couplings $y^{a}_{j}$ is free.
%In the case where the mass of the other fermions and scalars is nearly
%degenerated to $\psi_{1} $, the processes of coannihilation can be
%relevant for the relic density calculation~\cite{Klasen:2013jpa}.

@@

%\subsection{Scalar dark matter}
In contrast to fermion DM, the scalar DM candidate has additional interaction terms to the Yukawa and gauge interactions. This entails that the later ones can be used to alter the relic density predictions in scenarios with mixed scalar DM. Since in the present model the CP-even and CP-odd neutral $Z_2$-odd particles are mass degenerate, we have the scenario of singlet-doublet complex DM~\cite{Kadastik:2009dj,Belanger:2012vp} where the DM candidate is a mixture of a complex singlet \cite{McDonald:1993ex} and a $SU(2)_L$ doublet \cite{Deshpande:1977rw,Barbieri:2006dq}. It follows that for negligible Yukawa and gauge interactions there are two DM mass regions that allow us to properly reproduce the relic abundance, one corresponds the Higgs funnel region and the second one lies above $100$ GeV~\cite{Kakizaki:2016dza}.

%This corresponds to the simplified dark matter model of scalar singlet-doublet dark matter, which have been studied in detail in~\cite{Restrepo:2019soi}

%When the lightest scalar  is mainly a scalar doublet, the phenomenology will be similar inert doublet model~\cite{Honorez:2010re}. For this kind of DM model  inert doublet model there are two mass regions for DM which can to reproduce adequately the relic density ($50$GeV \textless $m_{DM}$ \textless $70$GeV and $535$GeV \textless $m_{DM}$ \textless $20$TeV)~\cite{Garcia-Cely:2015khw}.

@@

%Now, if the lightest scalar is mainly a singlet, the DM will be similar to Singlet %Scalar Dark Matter Model~\cite{Athron:2017kgt}\todo{change for the original ref by Ma et al. etc}.
%The mass range for the scalar singlet that produces the adequate relic
%density is around the Higgs resonance and for masses above $1$TeV. For
%this model~\cite{}, the relevant coupling for DM phenomenology is the
%Higgs portal trough the Lagrangian term $\sigma^{2} H^{2}_{i}$.
%This term generates interesting phenomenological consequences.
%Such as, thermal production of DM in the early
%universe~\cite{Yaguna:2008hd}, direct detection and invisible decay of
%SM Higgs through $h \to \sigma \sigma$~\cite{Mambrini:2011ik}.
%In addition, this model has important contributions for cosmology, in
%phenomena such as inflation~\cite{Lerner:2009xg} and
%baryogenesis~\cite{Cline:2012hg}.

\section{Cosmological constraints}
\label{sec:CosmoConstraints}
A not-so-minimal model can be obtained by introducing additional chiral SM singlet fields. 
In this case we can use the solutions to the anomaly cancellation conditions found for an Abelian gauge symmetry $\operatorname{U(1)_X}$ in~\cite{Jenkins:1987ue,Appelquist:2002mw,Campos:2017dgc,Okada:2018tgy,Jana:2019mez}.
We use $f$ ($f$) to denote the $X$-charge of the field $f_R$ ($F_L$) for one generation of the SM. The three linear anomalies in $\operatorname{U(1)_X}$ allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $\alpha$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \sum_{\alpha} n_{\alpha}+3 (e-2l)=&0\,, &   \sum_{\alpha} n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
There are many known solutions for the case in which $e-2l=1$. The full set of standard model charges in terms of $l$ is just
\begin{align}
  u=&-1-\frac{4l}{3}\,,&d=&1+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&1+2l\,,&h=&-1-l\,.
  \label{Eq:SMCharges}
\end{align}
The same solution but in terms of $h$ is given in ~\cite{Okada:2018tgy}.

@@

\section{Collider constraints}
ATLAS~\cite{}
recasting~\cite{}
In fig.~\ref{fig:u1blc}...

\begin{figure}
    \centering
    \includegraphics[scale=0.8]{u1blc}
    \caption{Constraints in the $\operatorname{U}(1)_{B-L}$ model.}
    \label{fig:u1blc}
\end{figure}

In the $\operatorname{U}(1)_{B-L}$, cosmlogical constraints are better that current ATLAS limit for $M_{Z'}\gtrsim 4.8\ \text{TeV}$

@@
